### 2024-05-10 example.tex
currently typos removed changing the figure and tabel labels

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Working Paper Template Example for
% IOE Graduate Conference 
% Version 15.0
%
% Original author:
% Jayandra Raj Shrestha (jayandra@ioe.edu.np)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% use the option blindreview to mask the author details for blind review submission
% remove the blinereview option to make the author information visible
\documentclass[fleqn, 10pt, twoside, blindreview]{IOEGC}

\hypersetup
{
    pdfauthor   = {author-1, author-2, author-3...}    % Edit with actual author names
}

\begin{document}
\maketitle
\thispagestyle{firstpage}

\section{Introduction}              
Nepal, a land of unparalleled biodiversity and stunning landscapes, is home to a richar-
ray of wildlife species that inhabit its lush forests, soaring mountains, and vibrante-
cosystems. The diverse wildlife, including endangered and elusive species like the
Bengal tiger, one-horned rhinoceros, and the elusive snow leopard, is a testament to the
country’s commitment to conservation. However, with the increasing challenges posed
by habitat loss, poaching, and climate change, preserving Nepal’s wildlife heritage de-
mands innovative and technologically advanced approaches. This project setsits sights
on creating a groundbreaking solution to bolster wildlife conservation efforts in Nepal
through an integrated approach for wildlife recognition and surveillance usingaudio and
video sensors. The synergy of audio and video data provides a comprehensive under-
standing of animal behavior, movement patterns, and species composition within their
natural habitats. By fusing these two streams of data, we aim to build arobust system
that can identify and monitor wildlife with a higher level of accuracy andefficiency.

\section{What is \LaTeX?} \label{sec:whatis}
\LaTeX{} is a document preparation system for the \TeX{}
typesetting program. It offers programmable desktop publishing features and 
extensive facilities for automating most aspects of typesetting and desktop  
publishing, including numbering and cross-referencing,
tables and figures, page layout, bibliographies, and much more.

\begin{itemize}[noitemsep]
	\item A family of programs designed to produce \\publication-quality typeset
		documents.
	\item Particularly good at working with mathematical symbols.
	\item WYSIWYM\footnote{What You See Is What You Mean} rather than 
		WYSIWYG\footnote{What You See Is What You Get}.
\end{itemize}

The history of LaTeX begins with a program called \TeX. In 1978, a computer 
scientist by the name of \textbf{Donald Knuth} grew frustrated with the mistakes
that his publishers made in typesetting his work. He decided to create a 
typesetting program that everyone could easily use to typeset documents, 
particularly those that include formulae, and made it freely available.

Knuth's product is an immensely powerful program, but one that does focus
very much on small details. A mathematician and computer scientist by the
name of Leslie Lamport wrote a variant of \TeX\ called \LaTeX\ that focuses on
document structure rather than such details.

\section{Getting the \LaTeX\ Software }\label{sec:getting}
There are two major standard distributions of \LaTeX:
\begin{itemize}[noitemsep]
	\item TeXLive \\ \texttt{\small{https://www.tug.org/texlive/}}
	\item MikTeX \\ \texttt{\small{https://miktex.org/}}
\end{itemize}

These are freely downloadable from the internet. TeXLive works in all the major 
PC platforms like Windows, Unix, Linux, and Mac. Whereas, MikTeX is for Windows 
only. When you install these, you also get the TeXWorks editor as your frontend.
More than a dozen other frontend GUIs are available for \LaTeX. 
Some of these are:

\begin{itemize}[noitemsep]
	\item TeXMaker \\ \texttt{\small{http://www.xm1math.net/texmaker/}}
	\item TeXnic Center\\ \texttt{\small{http://www.texniccenter.org/}}
\end{itemize}

\section{Template Structure} \label{sec:struct}
This \LaTeX\ template resides on a folder with the following files/folder:
\begin{description}
	\item[example.tex]		The main \LaTeX\ source file of this document (example.pdf). 
		Working Example on using the template with some description.
	\item[article.tex] A minimal alternate of example.tex.	
	\item[paperinfo.tex] Contains the abstract and other author information.
	\item[pagenum.tex]		Contains the code for starting page number which 
							will be edited during final compilation.
	\item[IOEGC.cls] 	\LaTeX\ class file for managing the styles and 
		formats of the document. Prohibited to edit.
	\item[refs.bib] 	File for placing the bibliography data in BibTeX format.
	\item[Graphics] 		Folder for keeping all the final graphics files 
							(.jpg, .png, etc.) used in the document.
\end{description}

\section{Sections}
Paragraphs within a document can be separated just by leaving one blank line 
between them.

\LaTeX\ supports section headings upto 3 levels via the following commands:

\begin{itemize}[noitemsep]
	\item \verb+\section{...}+
	\item \verb+\subsection{...}+
	\item \verb+\subsubsection{...}+
\end{itemize}

These have been illustrated properly in section \ref{sec:lists} of this example.
You can use their starred variants given below to suppress section numbering 
which has been demonstrated in the \emph{Acknowledgment} section.

\begin{itemize}[noitemsep]
	\item \verb+\section*{...}+
	\item \verb+\subsection*{...}+
	\item \verb+\subsubsection*{...}+
\end{itemize}

\section{Typesetting Mathematics}
\LaTeX\ has very rich features for typesetting mathematics. Please refer to 
\LaTeX\ and AMSmath manuals or online resources for further information. 
Here are a few examples.

The formula given in equation \ref{eq:quad} can be used to determine the roots 
of a quadratic equation of the form: $$ax^2+bx+c=0$$
Here, $a$, $b$, and $c$ are constants/coefficients and $x$ is a variable.

Numbered equation:
\begin{equation}
	x= \frac{-b\pm \sqrt{ b^2-4ac}}{2a}
	\label{eq:quad}
\end{equation}

Equation without a number
\begin{equation*}
	x = \frac{-b\pm \sqrt{ b^2-4ac}}{2a}
\end{equation*}

\section{Creating Tables}
Table \ref{tbl:capacity} is an example of a simple table in \LaTeX. To create 
complex tables, please refer to \LaTeX\ manuals or online resources. Use 
\verb+\begin{table*}+ to take up the entire page width. However, the use of 
tables spanning the entire page width is discouraged as it needs extra caution.

\begin{table}[H]
	\caption{No. of papers presented in IOEGC}
	\label{tbl:capacity}
	\centering
	\begin{tabular}{|c|l|c|} %three columns: left, center, center aligned
		\hline
		SN & Year & No. of papers \\
		\hline
		1 & 2013 & 42	\\
		2 & 2014 & 79	\\
		3 & 2015 & 46	\\
		4 & 2016 & 49	\\
		5 & 2017 & 83 	\\
		\hline
		 & Total & 299 	\\
		\hline
	\end{tabular}
\end{table}

\section{Placing Figures} \label{sec:figures}
One can generate technical graphs or diagrams from \LaTeX\ also, but this 
requires another level of expertise. Another alternate is to use R-programming 
code to generate graphs on the fly thus producing reproducible documents, 
which requires S-Weave. However, it is very common to include figures generated 
from other sources or programs. Here are a few examples on placing figures with 
proper captioning and label for cross referencing. The most suitable format for 
figure files to produce final output in raster format as pdf are:

\begin{enumerate}[noitemsep]
	\item PNG
	\item PDF
	\item JPG
\end{enumerate}

\begin{figure}[H]\centering
	% Remove fbox if you donot require a bounding box for the image
	\fbox{\includegraphics[width=0.95\linewidth]{curve}}
	\caption{Figure taking up 95\% width of the column}
	\label{fig:graph-1}
\end{figure}

% Using \begin{figure*} makes the figure take up the entire width of the page
% But be careful, the figure always comes at the top of the page ... 
% so it may appear on the next available page
\begin{figure*}[hbt]\centering
	\includegraphics[width=\linewidth]{prof_ale}
	\caption{Placing a wide picture (Discouraged! as it always appears at the 
			 top of a page.)}
	\label{fig:ale-sir}
\end{figure*}

Figure~\ref{fig:graph-1} takes up 95\% of the width of a column and 
Figure~\ref{fig:ale-sir} takes the width of the entire width of the page.

%forcing a column break
%\vfill\null

\section{Lists} \label{sec:lists}

\subsection{Simple Lists}
Simple Bulleted and Numbered lists have already been presented in 
Section~\ref{sec:getting} and Section~\ref{sec:figures} respectively.

\subsection{Nested Lists}
Lists can be nested upto three levels in \LaTeX.

\subsubsection{Numbered Nested List}
Here is a nested numbered list:

\begin{enumerate}
  \item Fruits
    \begin{enumerate}
      \item Apple
      \item Orange
    \end{enumerate}
  \item Vegetables
    \begin{enumerate}
      \item Spinach
      \item Carrot
    \end{enumerate}
\end{enumerate}

\subsubsection{Bulleted Nested List}
Here is a nested bulleted list:

% [noitemsep] removes whitespace between the items for a compact look
\begin{itemize}[noitemsep] 
  \item Fruits
    \begin{itemize}[noitemsep] 
      \item Apple
      \item Orange
    \end{itemize}
  \item Vegetables
    \begin{itemize}[noitemsep] 
      \item Spinach
      \item Carrot
    \end{itemize}
\end{itemize}

\subsubsection{Mixed Nested List}
Here is a mixed nested list:
\begin{enumerate}[noitemsep]
  \item Fruits
    \begin{itemize}[noitemsep] 
      \item Apple
      \item Orange
    \end{itemize}
  \item Vegetables
    \begin{itemize}[noitemsep] 
      \item Spinach
      \item Carrot
    \end{itemize}
\end{enumerate}

\subsection{Description List}
This is for dictionary-like word and description list.
\begin{description}
  \item[Word] Definition ...
  \item[Concept] Explanation ...
  \item[Idea] Text ...
\end{description}

\section{Paragraphs with heading}

\paragraph{Hello} Place your paragraph heading inside the curly braces and your 
paragraph text here.

\section{Referencing}
The list of references should be produced using BibTeX. The BibTeX entries 
should be placed in the "refs.bib" file. Please refer BibTeX manuals or online 
resources on creating bibliography databases using BibTeX and citation. You can 
easily create bibliography database files using the GUIs like TeXMaker or JabRef. 
You can even search for BibTeX entries for a majority of publications at Google 
Scholar in the following url:

\url{http://scholar.google.com}

Examples: This is citation one\cite{lamport1994} and these are two citations in 
one \cite{oetiker2001not, kopka1995guide}.

\section{Compilation} \label{sec:compile}
Since, this template contains citations and cross referencing along with 
reference list generated via BibTeX, the \LaTeX\ source file should be processed
four times in the following sequence to generate the final pdf output.
\begin{enumerate}
	\item PDFLatex
	\item BibTeX
	\item PDFLatex
	\item PDFLatex
\end{enumerate}

Do not worry, if there is an extra blank page at the end of the paper, this is 
an intended behavior. It happens to make the number of pages of the paper even, 
if the paper ends in an odd-numbered page. This is to make sure that every other
article always starts with an odd-numbered page.

\section{Submission} \label{sec:submit}
Before submitting the paper, the source file must be compiled without any error. 
The files that need to be submitted are:
\begin{itemize}[noitemsep]
\item article.tex (with your content)
\item article.pdf (with your content)
\item pagenum.tex (as-is)
\item paperinfo.tex (with your content)
\item refs.bib (with your content)
\item IOEGC.cls (as-is)
\item Graphics folder  (with your content)
\item Assets folder (with your content)
\end{itemize}
All these should be placed in compressed / zipped folder and submitted 
electronically.

\section{Review}
Your paper will be peer reviewed in blind by expert(s) before the conference. 
Comments may be provided in the submitted pdf file. You have to re-submit your 
paper by recompiling the \LaTeX\ source file as described in 
section~\ref{sec:compile} and submit as described in section \ref{sec:submit}.

\phantomsection
\section*{Still Having Problem?} 
\addcontentsline{toc}{section}{Problem?} 
There are a lot of online tutorials on \LaTeX\ available for free download. 
One of them being \textit{LaTeX Tutorials -- A Primer} by Indian \TeX\ Users 
Group \cite{ltxprimer}.

Further, there are websites like \url{sharelatex.com}, \url{overleaf.com}, etc.,
 which are very helpful
in finding out how to perform a specific task in \LaTeX\ .

If you still face technical problems in compiling your document in \LaTeX\ using
this template, please feel free to contact the primary author of this template 
via the following email address:

\centerline{\url{jayandra@ioe.edu.np}}

\section*{Future Enhancements}
\addcontentsline{toc}{section}{Future Enhancements} 
Lately, there has been a lot of demand for the creation of reproducible 
documents in research. One of the alternates in producing publication quality 
reproducible documents is the combination of \LaTeX\ and R-programming called 
S-weave.

In the near future, IOEGC is planning to adapt this mechanism to support 
reproducibility of research documents. Thus, you are highly encouraged to adapt 
this philosophy starting from this edition of IOEGC.

This template has undergone a few iterations of improvement over the past few 
years and is constantly evolving. Please feel free to send in your valuable 
comments/suggestions and/or feature requests via email to the primary author 
of this template.

\phantomsection
\section*{Acknowledgments} 
% The \section*{...} command prevents section numbering, and entry in TOC
% Add the following line to show this section in TOC
\addcontentsline{toc}{section}{Acknowledgments} 
The authors are thankful to ...

\phantomsection
\bibliographystyle{unsrt}
\putbib{refs}

\vfill\null
\end{document}

### 2024-05-10 paperinfo.tex
currently typos removed changing the figure and tabel labels

% Please do not leave any empty line before the closing braces (}).

\PaperTitle{An Integrated Approach for Wildlife Recognition in Nepal Using Video Along With Audio}

\Authors{
    Zishan Siddique \textsuperscript{a}, 
    Sandip Pariyar \textsuperscript{b}, 
    Venus Bastola\textsuperscript{c},
    Sushant Phagu \textsuperscript{d},
    Binay Lal Shrestha \textsuperscript{e},
    Pravin Sangroula \textsuperscript{f}
}

\Abstract{
Wildlife around the world is in decline primarily due to the loss of habitat as well as the intersection of territory between humans and wild animals. Manual recognition of animals
can be more accurate but will require exponentially greater resources both in terms of
capital and labor making it unfeasible in large-scale deployment, especially for a country like Nepal. We have developed a system for the recognition and classification of wild animals using deep convolutional neural network (DCNN)
architecture to aid conservation as well as as study of our ecological system. We used iNaturalist as our source for image data and Xeno-canto.org as the source for audio data. We were able to achieve an F1-Score of 86.12\% on image data of 44 species of animals and an F1-Score of 88.1\% on audio data of 23 species of birds all found in Nepal.
}

% If you don't have any keyword simply remove the text between the curly braces
% But DO NOT remove the \Keywords command
\Keywords{Convolutional Neural Network,EfficientNet, Audio, Video}

\Affiliation{
    \textsuperscript{a,b,c,d,e,f}
    \textit{Purwanchal Campus,IOE,Tribhuvan University,Nepal}
}
%\Affiliation{
  %  \textsuperscript{b}
  %  \textit{Affiliation of Author-2}

%\Affiliation{
   % \textsuperscript{c}
   % \textit{Affiliation of Author-3}
%}

\Email{
    \textsuperscript{a} 076bct095@ioepc.edu.np , 
    \textsuperscript{b} 076bct074@ioepc.edu.np ,
    \textsuperscript{c} 076bct091@ioepc.edu.np,
    \textsuperscript{d} 076bct090@ioepc.edu.np,
    \textsuperscript{d} binay@ioepc.edu.np,
    \textsuperscript{d} pravin@ioepc.edu.np

}

### 2024-05-10 article.tex
everything aligned right now but the cm,accuracy and loss for the audio model are now in the last page

% use the option blindreview to mask the author details for blind review submission
% remove the blinereview option to make the author information visible
\documentclass[fleqn, 10pt, twoside]{IOEGC}
\usepackage{listings}

\usepackage{svg}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{afterpage}
\usepackage{float}
\lstdefinestyle{mystyle}{
    language=Python,
    basicstyle=\small\ttfamily,,
    keywordstyle=\color{blue},
    commentstyle=\color{green},
    stringstyle=\color{purple},
    showstringspaces=false,
    numbers=left,
    numberstyle=\tiny,
    numbersep=5pt,
    breaklines=true,
    breakatwhitespace=true,
    tabsize=2
}
\hypersetup
{
    pdfauthor   = {Zishan Siddique, Sandip Pariyar, Venus Bastola, Sushant Phagu, Binay lal Shrestha, Pravin Sangroula},	% Edit with actual author names 
}

\begin{document}

\maketitle
\thispagestyle{firstpage}

\section{Introdution}
Nepal is a land of unparalleled biodiversity with 208 species of mammals \cite{r1}, 142 species of reptiles \cite{r2}, and more than 873 species of birds \cite{r3}. The diverse wildlife, including endangered species found here, are testament to the country’s commitment to conservation. However, with the increasing challenges posed
by habitat loss, poaching, and climate change, preserving Nepal’s wildlife heritage demands innovative and technologically advanced approaches.
This project provides a basic template for creating solutions to bolster wildlife conservation efforts in Nepal.
\par
Vision can be sufficient for some species that are of adequate size but vision alone can miss many opportunities for identification and classification. For this reason, we have used two models i.e. vision and audio for a broader scope.
The paper is structured as follows:
\begin{itemize}
	\item Earlier approaches for our objectives are discussed in Section 2.
	\item The datasets used and their metainformation are explained in Section 3.
	\item Our approach for the proposed system is elaborated in Section 4.
	\item The results of our solution are explained in Section 5.
	\item Finally, the complete project is concluded with a brief paragraph in Section 6.
\end{itemize}

\section{Background and Related Works}
There have been many instances of work done for wildlife. In 2017, Nguyen et al.\cite{r4} worked on a monitoring system for animals in South-central Victoria, Australia with appreciable results. HOW?
\par
Gautam et al. 2023 \cite{r5} worked on the classification of birds based on audio data of 44 species found in Nepal. Their work was based on Mel Spectogram features which are often termed "pictures of a sound". Mel Spectogram and MFCC (Mel-Frequency Cepstral Coefficients) were used for feature extraction of audio data which was used to train a Convolutional Neural Network.
\par
In 2021, Qi et al. \cite{r6} worked on image recognition in Animal Husbandry. They applied super pixel-based image segmentation and SIFT algorithm to complete image segmentation and feature extraction which was later modeled using Convolutional Neural Network and SVM for classification.
\par
We propose an integrated system of using both audio and video. The audio is used as it is with some pre-processing and the video is broken into frames which are fed to the image classifier model.

\section{Dataset Analysis}
As two mediums vision and audio are being used for classification of animals, two different datasets comprising of labeled images and labeled audio are required.
The image dataset has a total of 44 classes. Of the 44 classes, there are 31 mammals, 9 birds, and 4 reptiles. Moreover, we couldn’t find research-grade audio data for groups of animals other than birds. Thus, the audio model comprises 23 classes of birds all found in Nepal.
\subsection{Image Data from iNaturalist}
iNaturalist is an online social network of people sharing biodiversity information. It has a large size of data which includes images, audio, and video of large number of species found around the world. The data is collected by users who submit it to the site under the public domain, Creative Commons, or with all rights reserved. iNaturalist encourages its users to opt for more open licenses like Creative Commons for the research community.We collected a total of 38,442 images spanning across 44 species all found in Nepal. As the data contained some anomalous images, we had to manually remove them. After the cleaning step, the data was preprocessed as per the Efficient Net Model for better training. Of the 38,442 images collected, 50 images from each class were used as test datasets.
API provided by iNaturalist was used for downloading the data. The bar chart in Figure \ref{image-file-count-animal} shows the frequency of each class of image.

\begin{figure}
	\centering
	\includegraphics[scale=0.5]{Graphics/naturalist_horizontal.png}
	\caption{Image File count for each animal}
	\label{image-file-count-animal}
\end{figure}

\subsection{Audio Data from Xeno-canto}
Xeno-canto is a website dedicated to sharing wildlife sounds from all over the world. It provides its data under several licenses derived from Creative Commons licenses which are free for use for educational and research purposes. We collected a total of thousands of audio spanning 23 classes and performed temporal segmentation with each segment being 5 seconds long. A total of 16,230 audio files each of 5 sec was now our final dataset. Of the 16,230 audio files, 30 files for each class were used as the test dataset.
API provided by Xeno-canto was used for downloading the data. The bar chart in Figure \ref{audio-file-count-bird} shows the frequency of each class of audio.

	{

		\begin{figure}
			\centering
			\includegraphics[scale=0.5]{Graphics/xenocanto.png}
			\caption{Audio File count for each bird}
			\label{audio-file-count-bird}
		\end{figure}

	}
\section{Proposed Methodology}
\subsection{Mel-frequency cepstrum Coefficients based on DCT-II for Audio Feature Extraction
}
Mel-frequency cepstrum (MFC) refers to the short-term power spectrum of a sound
based on cosine transform. Mel-frequency cepstrum Coefficients(MFCCs) are coefficients that collectively make up an MFC. Mel-frequency cepstrum Coefficients transform is done on the audio to extract a small set of features that describes our audio.
MFCC helps us to capture the essential spectral characteristics of speech signals in a
compact form, making them suitable for pattern recognition tasks.
\begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
def mfcc_transform(y: np.ndarray,
                   sr: int,
                   mfcc_params: Dict
                   ) -> torch.Tensor:
    melspec = librosa.feature.melspectrogram(
        y=y,
        sr=sr,
        n_fft=mfcc_params['n_fft'],
        win_length=mfcc_params['win_length'],
        hop_length=mfcc_params['hop_length'],
        n_mels=mfcc_params['n_mels'],
        htk=mfcc_params['htk'],
        norm=mfcc_params['norm_melspec'],
    )
    mfcc_librosa = librosa.feature.mfcc(
        S=librosa.core.spectrum.power_to_db(melspec),
        n_mfcc=mfcc_params['n_mfcc'],
        dct_type=mfcc_params['dct_type'],
        norm=mfcc_params['norm_mfcc']
    )

    return torch.from_numpy(mfcc_librosa)
\end{lstlisting}

\subsection{Expansion and Interpolation of Audio Data}
After MFCC transformation, the tensors are transformed again to proper dimension as required by EfficientNetB0 architecture. Following steps are performed:
\begin{itemize}
	\item The input tensor y is replicated along one dimension to create three copies, preparing it for subsequent operations.
	\item The expanded tensor is resized to a fixed size of (224, 224) using bilinear interpolation, ensuring consistent dimensions for neural network input.
	\item Any extra dimensions introduced during resizing are removed, returning the tensor to its original shape.
	\item The processed tensor is returned for further use, completing the preprocessing steps commonly applied to image data for deep learning tasks.
\end{itemize}
\begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
def expand_tensor_for_efficient_net(y) -> torch.Tensor:
    input_tensor_expanded = y.expand(3, -1, -1)
    resized_tensor = F.interpolate(input_tensor_expanded.unsqueeze(
        0), size=(224, 224), mode='bilinear', align_corners=False)

    resized_tensor = resized_tensor.squeeze()
    return resized_tensor
\end{lstlisting}

\subsection{Transformation as required by EfficientNetB0 Input Data}
For best results and compatibility, we transformed our data as per the transformation required for EfficientNetB0 architecture.
\begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
ImageClassification(
    crop_size=[224]
    resize_size=[256]
    mean=[0.485, 0.456, 0.406]
    std=[0.229, 0.224, 0.225]
    interpolation=InterpolationMode.BICUBIC
)
\end{lstlisting}

\subsection{Transfer Learning}
Transfer learning is a technique in the field of machine learning where learned parameters on previously trained data are used to learn parameters of new data. This is done to reuse the learned parameters from previous data. Transfer learning is not
only better for performance but also decreases the time required to learn features about
subject in question.

\subsection{EfficientNetB0}
EfficientNet is a convolutional neural network architecture and scaling method that uniformly scales all dimensions of depth/width/resolution using a compound coefficient \cite{r7}.
Unlike conventional practice that arbitrarily scales these factors, the EfficientNet scaling method uniformly scales network width, depth, and resolution with a set of fixed scaling
coefficients.
\par
\subsection{ Convolutional Neural Network}
Convolutional Neural Networks are specialized kind of neural networks for processing
data that has a grid-like topology \cite{r8}. Most machine learning libraries implement cross-correlation rather than conventionally known convolution operations in the realm of Digital
Signal Processing . In conventional multi-layer perceptrons, large number of parameters are required to learn features where as CNN uses filters to stride through whole data using a lesser number of parameters aka parameter sharing.
% CNNs are often applied in large number of computer vision tasks across the industry.

\section{Result and Analysis}

\begin{center}
	\captionof{table}{Overview of Results of Audio and Video Model}
	\begin{tabular}{ |c|c|c| }
		\hline
		Model  & Accuracy & F1 score \\
		\hline
		Vision & 86.1     & 86.12    \\
		\hline
		Audio  & 90.39    & 88.12    \\
		\hline
	\end{tabular}
\end{center}

\subsection{Vision Model}
The vision model was created with EfficientNetB0 as its base for transfer learning. We were able to achieve an accuracy of 86.1\% with an F1-Score of 86.12 on the test dataset. Table \ref{table1} shows the various metrics of the vision model.
\begin{center}
	\captionof{table}{Vision Model Metrics}
	\begin{tabular}{ |c|c|c| }
		\hline
		\textbf{Parameters}      & \textbf{Metrics}          \\
		\hline
		Batch Size               & 32                        \\
		\hline
		Epochs                   & 10                        \\
		\hline
		Pretrained Model         & EfficientNetB0            \\
		\hline
		Weights                  & ImageNet                  \\
		\hline
		Learning Rate            & 0.0001                    \\
		\hline
		Weight Decay             & 0.001                     \\
		\hline
		Loss Function            & Categorical Cross Entropy \\
		\hline
		Optimizer                & Adam                      \\
		\hline
		Total Parameters         & 4,063,912                 \\
		\hline
		Trainable Parameters     & 4,063,912                 \\
		\hline
		Non-Trainable Parameters & 0                         \\
		\hline
	\end{tabular}
	\label{table1}

\end{center}

The confusion matrix for vision model on the test dataset is shown in Figure \ref{cm_vision}. Figure \ref{accuracy-vision} and Figure \ref{loss-vision} shows the accuracy and loss of vision model respectively.

\begin{figure*}[t] % Use figure* for spanning both columns
	\centering
	\begin{minipage}{\textwidth} % Use minipage to ensure the figure spans the whole width
		\includegraphics[scale=0.53]{Graphics/vision_cm.png} % Replace example-image with your actual image file
		\caption{Confusion Matrix for Vision Model on test data}
		\label{cm_vision}
	\end{minipage}
\end{figure*}

\begin{figure*}[hbt]
	\centering
	\begin{minipage}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/accuracy_vision_model.png}
		\caption{Accuracy of Vision Model on Train and Test Data}
		\label{accuracy-vision}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/loss_vision.png}
		\caption{Loss of Vision Model on Train and Test Data}
		\label{loss-vision}
	\end{minipage}
\end{figure*}

\clearpage

\subsection{Audio Model}

The audio model was also created with EfficientNetB0 as its base for transfer learning after feature extraction done with MFCC Transform which was again transformed as explained in Section 4.3 for input compatibility with EfficientNetB0.
\par
We were able to achieve an accuracy of 90.39\% and an F1-Score of 88.1 on the test dataset. The batch size of 64 was chosen to increase the training pace. The size of test size of each class is 30.
\par
The following table shows various metrics of the audio model:

\begin{center}
	\captionof{table}{Audio Model Metrics}
	\begin{tabular}{ |c|c|c| }
		\hline
		\textbf{Parameters}      & \textbf{Metrics}          \\
		\hline
		Batch Size               & 64                        \\
		\hline
		Epochs                   & 15                        \\
		\hline
		Pretrained Model         & EfficientNetB0            \\
		\hline
		Weights                  & ImageNet                  \\
		\hline
		Learning Rate            & 0.0001                    \\
		\hline
		Weight Decay             & 0                         \\
		\hline
		Loss Function            & Categorical Cross Entropy \\
		\hline
		Optimizer                & Adam                      \\
		\hline
		Total Parameters         & 4,037,011                 \\
		\hline
		Trainable Parameters     & 4,037,011                 \\
		\hline
		Non-Trainable Parameters & 0                         \\
		\hline
	\end{tabular}
\end{center}
We see a similar pattern as seen in the vision model. Here too in the audio model with similar featured data being mislabeled. Figure \ref{accuracy-audio} and Figure \ref{loss-audio} shows the accuracy and loss of audio model respectively on test data. Figure \ref{cm_audio} shows confusion matrix for audio model on test dataset.

\begin{figure*}[hbt]
	\centering
	\begin{minipage}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/audio_accuracy.png}
		\caption{Accuracy of Audio Model on Train and Test Data}
		\label{accuracy-audio}
	\end{minipage}
	\hfill
	\begin{minipage}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/audio_loss.png}
		\caption{Loss of Audio Model on Train and Test Data}
		\label{loss-audio}
	\end{minipage}
\end{figure*}

\begin{figure*}[t] % Use figure* for spanning both columns
	\centering
	\begin{minipage}{\textwidth} % Use minipage to ensure the figure spans the whole width
		\includegraphics[scale=0.4]{Graphics/audio_cm.png} % Replace example-image with your actual image file
		\caption{Confusion Matrix for Audio Model on test data}
		\label{cm_audio}
	\end{minipage}
\end{figure*}

\section{Conclusion}
Our research work in wildlife classification combines audio and visual technologies, showing promising accuracy levels of 90.39\% and 86.1\% for audio and vision models, respectively. Despite challenges in distinguishing similar species, our study represents a significant advancement in leveraging technology for wildlife conservation. By enhancing monitoring capabilities, we aim to contribute to sustainable conservation practices and promote harmonious coexistence between humans and wildlife.

\phantomsection
\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}
The authors would like to express their gratitude to everyone who supported
them throughout this project. The authors are thankful for their aspiring guidance,
invaluably constructive criticism, and friendly advice during the project work. The authors are
grateful to them for sharing their views on issues related to the project. Without the aid
of our friends and family members, our project would not have been attainable. This
feat would not have been achieved without their assistance.

\phantomsection

\bibliographystyle{unsrt}
\bibliography{references}

\vfill\null
\end{document}

### 2024-05-10 article.tex
everything fixed

@@
 	\end{tabular}
+	\label{table:overview}
 \end{center}
@@
 We were able to achieve an accuracy of 86.1\% with an F1-Score of 86.12 on the test dataset.
-Table \ref{table1} shows the various metrics of the vision model.
+Table \ref{table:vision-model} shows the various metrics of the vision model.
 \begin{center}
@@
 	\end{tabular}
-	\label{table1}
+	\label{table:vision-model}
 
@@
 
+Figure \ref{accuracy-vision} and Figure \ref{loss-vision} shows the accuracy and loss of vision model respectively.
 The confusion matrix for vision model on the test dataset is shown in Figure \ref{cm_vision}.
-Figure \ref{accuracy-vision} and Figure \ref{loss-vision} shows the accuracy and loss of vision model respectively.
-
-\begin{figure*}[t] % Use figure* for spanning both columns
-	\centering
-	\begin{minipage}{\textwidth} % Use minipage to ensure the figure spans the whole width
-		\includegraphics[scale=0.53]{Graphics/vision_cm.png} % Replace example-image with your actual image file
-		\caption{Confusion Matrix for Vision Model on test data}
-		\label{cm_vision}
-	\end{minipage}
-\end{figure*}
 
@@
 
+\begin{figure*}[t] % Use figure* for spanning both columns
+	\centering
+	\begin{minipage}{\textwidth} % Use minipage to ensure the figure spans the whole width
+		\includegraphics[scale=0.53]{Graphics/vision_cm.png} % Replace example-image with your actual image file
+		\caption{Confusion Matrix for Vision Model on test data}
+		\label{cm_vision}
+	\end{minipage}
+\end{figure*}
+
 \clearpage
@@
 \par
-We were able to achieve an accuracy of 90.39\% and an F1-Score of 88.1 on the test dataset. The batch size of 64 was chosen to increase the training pace. The size of test size of each class is 30.
-\par
-The following table shows various metrics of the audio model:
+We were able to achieve an accuracy of 90.39\% and an F1-Score of 88.1 on the test dataset.
+The size of test size of each class is 30.
+Table \ref{table:audio-model} shows various metrics of the audio model:
 
@@
 	\end{tabular}
+	\label{table:audio-model}
 \end{center}
@@
 Here too in the audio model with similar featured data being mislabeled.
-Figure \ref{accuracy-audio} and Figure \ref{loss-audio} shows the accuracy and loss of audio model respectively on test data.
+Figure \ref{accuracy-audio} and Figure \ref{loss-audio} shows the accuracy and loss of audio model respectively.
 Figure \ref{cm_audio} shows confusion matrix for audio model on test dataset.
-
-\begin{figure*}[hbt]
+\begin{figure*}[ht]
 	\centering
-	\begin{minipage}[t]{0.45\textwidth}
+	\begin{minipage}[ht]{0.45\textwidth}
 		\centering
@@
 	\hfill
-	\begin{minipage}[t]{0.45\textwidth}
+	\begin{minipage}[ht]{0.45\textwidth}
 		\centering
@@
 
+\clearpage
 \section{Conclusion}

### 2024-05-10 article.tex
final spelling mistakes solved

@@
     language=Python,
-    basicstyle=\small\ttfamily,,
+    basicstyle=\footnotesize,
     keywordstyle=\color{blue},
@@
     breakatwhitespace=true,
-    tabsize=2
+    tabsize=1
 }
@@
 	\item Earlier approaches for our objectives are discussed in Section 2.
-	\item The datasets used and their metainformation are explained in Section 3.
+	\item The datasets used and their meta information are explained in Section 3.
 	\item Our approach for the proposed system is elaborated in Section 4.
@@
 \section{Background and Related Works}
-There have been many instances of work done for wildlife. In 2017, Nguyen et al.\cite{r4} worked on a monitoring system for animals in South-central Victoria, Australia with appreciable results. HOW?
+There have been many instances of work done for wildlife.
+In 2017, Nguyen et al.\cite{r4} worked on a monitoring system for animals in South-central Victoria, Australia with appreciable results.
+They used deep convolutional neural networks (DCNNs) to train a model on 3.2 million images of animals spanning 48 species from Tanzania's Serengeti National Park.
 \par
@@
 \section{Dataset Analysis}
-As two mediums vision and audio are being used for classification of animals, two different datasets comprising of labeled images and labeled audio are required.
+Two mediums vision and audio are being used for
+classification of animals, two different datasets comprising of
+labeled images and labeled audio are required. ImageNet was our
+the initial choice for image data but due to the misrepresentation of
+species, the idea was dropped, and iNaturalist was chosen as our
+source for image data for its reliability and proper
+representation\cite{r9}.
 The image dataset has a total of 44 classes.
@@
 The data is collected by users who submit it to the site under the public domain, Creative Commons, or with all rights reserved.
-iNaturalist encourages its users to opt for more open licenses like Creative Commons for the research community.We collected a total of 38,442 images spanning across 44 species all found in Nepal.
+iNaturalist encourages its users to opt for more open licenses like Creative Commons for the research community.
+We collected a total of 38,442 images spanning 44 species all found in Nepal.
 As the data contained some anomalous images, we had to manually remove them.
@@
 Mel-frequency cepstrum Coefficients transform is done on the audio to extract a small set of features that describes our audio.
-MFCC helps us to capture the essential spectral characteristics of speech signals in a
+MFCC helps us capture the essential spectral characteristics of speech signals in a
 compact form, making them suitable for pattern recognition tasks.
+% default_mfcc_params = {
+%     "n_fft": 2048,
+%     "win_length": None,
+%     "n_mels": 256,
+%     "hop_length": 512,
+%     "htk": True,
+%     "norm_melspec": None,
+%     "n_mfcc": 256,
+%     "norm_mfcc": "ortho",
+%     "dct_type": 2,
+% }
 \begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
 def mfcc_transform(y: np.ndarray,
-                   sr: int,
-                   mfcc_params: Dict
+                   sr: int
                    ) -> torch.Tensor:
@@
         sr=sr,
-        n_fft=mfcc_params['n_fft'],
-        win_length=mfcc_params['win_length'],
-        hop_length=mfcc_params['hop_length'],
-        n_mels=mfcc_params['n_mels'],
-        htk=mfcc_params['htk'],
-        norm=mfcc_params['norm_melspec'],
+        n_fft=2048,
+        win_length=2048,
+        hop_length=512,
+        n_mels=256,
+        htk=True,
+        norm="ortho",
     )
@@
         S=librosa.core.spectrum.power_to_db(melspec),
-        n_mfcc=mfcc_params['n_mfcc'],
-        dct_type=mfcc_params['dct_type'],
-        norm=mfcc_params['norm_mfcc']
+        n_mfcc=256,
+        dct_type=2,
+        norm="ortho"
     )
@@
 \subsection{Expansion and Interpolation of Audio Data}
-After MFCC transformation, the tensors are transformed again to proper dimension as required by EfficientNetB0 architecture. Following steps are performed:
+After MFCC transformation, the tensors are transformed again to the proper dimension as required by EfficientNetB0 architecture. The following steps are performed:
 \begin{itemize}
@@
 data that has a grid-like topology \cite{r8}. Most machine learning libraries implement cross-correlation rather than conventionally known convolution operations in the realm of Digital
-Signal Processing .
+Signal Processing.
 In conventional multi-layer perceptrons, large number of parameters are required to learn features where as CNN uses filters to stride through whole data using a lesser number of parameters aka parameter sharing.
@@
 	\captionof{table}{Overview of Results of Audio and Video Model}
-	\begin{tabular}{ |c|c|c| }
+	\begin{tabular}{ |c|c|c|c| }
 		\hline
-		Model  & Accuracy & F1 score \\
+		Model  & Train Accuracy & Test Accuracy & F1 score on Test Data \\
 		\hline
-		Vision & 86.1     & 86.12    \\
+		Vision & 96.00          & 86.10         & 86.12                 \\
 		\hline
-		Audio  & 90.39    & 88.12    \\
+		Audio  & 98.36          & 90.39         & 88.12                 \\
+
 		\hline
@@
 The vision model was created with EfficientNetB0 as its base for transfer learning.
-We were able to achieve an accuracy of 86.1\% with an F1-Score of 86.12 on the test dataset.
+We were able to achieve an accuracy of 86.12\% with an F1-Score of 86.12 on the test dataset.
 Table \ref{table:vision-model} shows the various metrics of the vision model.
@@
 
-Figure \ref{accuracy-vision} and Figure \ref{loss-vision} shows the accuracy and loss of vision model respectively.
-The confusion matrix for vision model on the test dataset is shown in Figure \ref{cm_vision}.
+Figure \ref{accuracy-vision} and Figure \ref{loss-vision} show the accuracy and loss of vision model respectively.
+The confusion matrix for the vision model on the test dataset is shown in Figure \ref{cm_vision}.
 
@@
 		\includegraphics[scale=0.53]{Graphics/vision_cm.png} % Replace example-image with your actual image file
-		\caption{Confusion Matrix for Vision Model on test data}
+		\caption{Confusion Matrix for Vision Model on Test Data}
 		\label{cm_vision}
@@
 \par
-We were able to achieve an accuracy of 90.39\% and an F1-Score of 88.1 on the test dataset.
-The size of test size of each class is 30.
-Table \ref{table:audio-model} shows various metrics of the audio model:
+We were able to achieve an accuracy of 90.39\% and an F1-Score of 88.12 on the test dataset. The test size of each class is 30. Table \ref{table:audio-model} shows various metrics of the audio model:
 
@@
 We see a similar pattern as seen in the vision model.
-Here too in the audio model with similar featured data being mislabeled.
-Figure \ref{accuracy-audio} and Figure \ref{loss-audio} shows the accuracy and loss of audio model respectively.
-Figure \ref{cm_audio} shows confusion matrix for audio model on test dataset.
+Here too is the audio model with similar featured data being mislabeled.
+Figure \ref{accuracy-audio} and Figure \ref{loss-audio} show the accuracy and loss of the audio model respectively.
+Figure \ref{cm_audio} shows the confusion matrix for the audio model on the test data.
 \begin{figure*}[ht]
@@
 		\includegraphics[scale=0.4]{Graphics/audio_cm.png} % Replace example-image with your actual image file
-		\caption{Confusion Matrix for Audio Model on test data}
+		\caption{Confusion Matrix for Audio Model on Test Data}
 		\label{cm_audio}

### 2024-05-10 paperinfo.tex
final spelling mistakes solved

@@
 We have developed a system for the recognition and classification of wild animals using deep convolutional neural network (DCNN)
-architecture to aid conservation as well as as study of our ecological system.
+	architecture to aid conservation as well as a study of our ecological system.
 We used iNaturalist as our source for image data and Xeno-canto.org as the source for audio data.

### 2024-05-10 article.tex
final fixes done

@@
 \section{Dataset Analysis}
-Two mediums vision and audio are being used for
+As two mediums, vision and audio are being used for
 classification of animals, two different datasets comprising of
@@
 We see a similar pattern as seen in the vision model.
-Here too is the audio model with similar featured data being mislabeled.
+Here too, the audio model is struggling to classify audio of similar species.
 Figure \ref{accuracy-audio} and Figure \ref{loss-audio} show the accuracy and loss of the audio model respectively.
@@
 \section{Conclusion}
-Our research work in wildlife classification combines audio and visual technologies, showing promising accuracy levels of 90.39\% and 86.1\% for audio and vision models, respectively.
+Our research work in wildlife classification combines audio and visual technologies, showing promising accuracy levels of 90.39\% and 86.1\% for audio and vision models on test data, respectively.
+Moreover, the audio and vision model achieves an F1 score of 86.12 and 88.12 on test data respectively.
 Despite challenges in distinguishing similar species, our study represents a significant advancement in leveraging technology for wildlife conservation.
